### 2011-05-03 thesis.tex
Should have done this earlier...

\documentclass[10pt,a4paper]{report}
\usepackage[latin1]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\author{James Pettit}
\begin{document}
\section{Introduction}
\section{The Game of Hex}
\section{State of the Art in Monte Carlo Tree Search}
UCT
AMAF
RAVE
Heuristics
Tree Bias
\section{Problem Statement}
\section{Solution}
\section{Conclusion}
\subsection{Conclusions}
Conclusions are not a rambling summary of the thesis: they are short, concise statements of the inferences that you have made because of your work. It helps to organize these as short numbered paragraphs, ordered from most to least important.
\subsection{Summary of Contributions}
The Summary of Contributions will be much sought and carefully read by the examiners. Here you list the contributions of new knowledge that your thesis makes. Of course, the thesis itself must substantiate any claims made here. There is often some overlap with the Conclusions, but that's okay. Concise numbered paragraphs are again best. Organize from most to least important.
\subsection{Future Research}
The Future Research subsection is included so that researchers picking up this work in future have the benefit of the ideas that you generated while you were working on the project. Again, concise numbered paragraphs are usually best.
\end{document}