### 2012-05-09 presentation/presentation.tex
More stuff to talk about.

\documentclass{beamer}
\usetheme{Boadilla}
\usepackage{havannah}

\title[Learning MCTS]{Evolutionary Learning of Policies for MCTS Simulations}
\author[Pettit, Helmbold]{James Pettit, David Helmbold}
\institute[UCSC]{
  University of California, Santa Cruz\\
  \texttt{jpettit@soe.ucsc.edu}
}
\date[July 2012]{July 2012}

\begin{document}

\begin{frame}[plain]
  \titlepage
\end{frame}

\begin{frame}{Overview}
\begin{enumerate}
	\item The Game of Hex
	\item Monte-Carlo Tree Search (MCTS)
	\item Apply Evolutionary Learningto MCTS+Hex
	\item Results and Future Work
\end{enumerate}
\end{frame}

\begin{frame}{The Game of Hex}
\begin{itemize}
	\item 2 player, perfect information
	\item 6-sided hexagons on a parallelogram board
	\item Common sizes: 11, 13
\end{itemize}
\end{frame}

\begin{frame}{The Game of Hex - Example Board}
% example board, play a few moves, show final position
	\begin{figure}[tb]
	\resizebox{3.3in}{!}{ \begin{HexBoard}[board size=7]
		  \HGame{d4,d3,e3,f1,c2,d6,b5,e2,c4,c3,a4,b6,c6,c5,d5,b7,c7,b4,a5,b3,a3,b1,b2,c1,d1}
		\end{HexBoard}
		}
	\end{figure}
\end{frame}

\begin{frame}{The Game of Hex - Good for AI}
\begin{itemize}
	\item Easy to program
	\item Clear-cut winning condition
	\item Large problem space
	\item Solved for boards up to 7x7
\end{itemize}
\end{frame}

\begin{frame}{Tree Search}
\begin{itemize}
	\item Game tree grows exponentially
	\item Symmetry can halve space
	\item Limited opportunities for provable pruning
	\item No good position ranking heuristic
\end{itemize}
\end{frame}

\begin{frame}{Monte Carlo Tree Search}
\begin{itemize}
	\item Use random playouts to estimate minimax value
	\item Large enough playouts will converge to true minimax value
	\item Seems dumb, actually works
	\item Computationally very expensive
\end{itemize}
\end{frame}

\begin{frame}{Monte Carlo Tree Search - Playout Policy}
\begin{itemize}
	\item Naively ``improving'' the strength of the playout policy can hurt overall performance
	\item Requires careful and expensive testing to verify improvement
\end{itemize}
\end{frame}

\begin{frame}{Evolutionary Learning}
\begin{itemize}
	\item Idea: Evolve playout policy
	\item Individual policies compete in a tournament to reproduce
	\item Self-play yields a self-bootstrapping system
\end{itemize}
\end{frame}

\begin{frame}{Encoding}
\begin{itemize}
	\item Individual policy is a collection of evolvable weights
	\item One weight, many interpretations
	\item Explicitly requires domain knowledge
	\item Implicitly limits the system
\end{itemize}
\end{frame}

\begin{frame}{Results}
\begin{itemize}
	\item Strong and robust
\end{itemize}
\end{frame}

\begin{frame}{Future Work}
\begin{itemize}
	\item Different encodings
	\item General game play
	\item Encode expert knowledge in weights
\end{itemize}
\end{frame}

\end{document}